Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-05-14 writing/embedded_meta_learning.tex
Rest of Erin's comments

\begin{abstract}
How can deep learning systems flexibly reuse their knowledge? Toward this goal, we propose a new class of challenges, and a class of architectures that can solve them. The challenges are in the form of meta-mappings, which involve systematically transforming task behaviors to adapt to new tasks. We suggest that the key to achieving these challenges is representing the task being performed along with the computations used to perform it. We therefore draw inspiration from meta-learning and functional programming to propose a class of Embedded Meta-Learning (EML) architectures that represent both data and tasks in a shared latent space. We demonstrate the flexibility of these architectures by showing that they can perform meta-mappings, i.e. that they can exhibit zero-shot remapping of behavior on new tasks. 
\end{abstract}

@@

\section{Embedded Meta-Learning (EML) architecture}
To achieve this, we propose a new class of Embedded Meta-Learning (EML) architectures which consist of two parts: 
\vspace{-0.75em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to perform tasks in accordance with their embeddings.
\end{enumerate}
\vspace{-0.75em}
The utility of having a completely shared space $Z$ in which data, language, and tasks are represented is that it allows for arbitrary mappings between these distinct types of data. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explaining behavior), or tasks to tasks (changing behavior based on examples). This is a step closer to the flexible reasoning humans display. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize these to new instances drawn from the same data/task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow. \par 
\begin{figure}
\centering
\begin{tikzpicture}[auto, scale=0.75]
\node [dashblock] at (0, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: \textbf{Shared} \\ \textbf{representation}\end{tabular}};

@@

\node [conc] at (-7, -4) (perc) {\textbf{Input}};
%%\node [conc] at (-7.25, -2.5) (lang) {\textbf{Language}};
\node [conc] at (-7, 4) (act) {\textbf{Output}};
\node [block, text width=2cm] at (-4.5, -2) (IE) {$\bm{\mathcal{I}}$: \textbf{Input Encoder}};
%%\node [block, text width=2cm] at (-5, -0.5) (LE) {$\bm{\mathcal{L}}$: \textbf{Lang. Encoder}};
\node [block, text width=2.2cm] at (-4, 2.5) (OD) {$\bm{\mathcal{O}}$: \textbf{Output Decoder}};
\node [block, text width=2cm] at (-5.5, 1) (TE) {$\bm{\mathcal{T}}$: \textbf{Target Encoder}};
\path [line] (perc.north) to (IE.south);
%%\path [line] (lang.north) to ([xshift=0.05cm, yshift=0.05cm]LE.south west);
\path [line] ([xshift=-0.05cm, yshift=-0.05cm]IE.north east) to (rep.south west);
%%\path [line] (LE.east) to (rep.west);
\path [line] (rep.north west) to ([xshift=-0.05cm, yshift=0.05cm]OD.south east);
\path [line] (OD.north) to (act.south);
\path [line] (act.south) to (TE.north);
\path [line] (TE.east) to (rep.north west);

@@

Specifically, we embed inputs, targets, and mappings into a shared representational space $Z$. Inputs are embedded by a deep network $\mathcal{I}: \text{input} \rightarrow Z$. Outputs are decoded from the representational space by a deep network $\mathcal{O}: Z \rightarrow \text{output}$. Targets are encoded by a deep network $\mathcal{T}: \text{targets} \rightarrow Z$. \par
Given this, the task of mapping inputs to outputs can be framed as trying to find a transformation of the representational space that takes the (embedded) inputs from the training set to the (embedded) targets. These transformations are performed by a system with the following components (see Fig. \ref{architecture_fig} for a schematic):  

@@

To write this explicitly, suppose we have some dataset of input, target pairs ($D_1 = \{(x_0, y_0), ...\}$), and some input $x$ for which we wish to generate a predicted output $\hat{y}$. This output would be generated as follows: 

@@

This system can be trained end-to-end if labels are provided for the second set of inputs. In particular, suppose we have some loss function $\mathcal{L}(y, \hat{y})$ defined on a single target output $y$ and actual model output $\hat{y}$, for some input $x$. We define our total loss computed on some dataset $D_2$ as:

@@

More generally, suppose we have some input which is already embedded in the representation space $z_{in} \in Z$, and a task/function embedding $z_{\text{function}} \in Z$. Then we can generate an output $\hat{z}_{out} \in Z$ as:  
\[\hat{z}_{out} = f_{z_{\text{function}}}(z_{in}) \qquad \text{where } f_{z_{\text{function}}} \text{ is parameterized by } \mathcal{H}\left(z_{\text{function}}\right)\]
This output can then be appropriately dispatched depending on the task at hand. For example, if the $z_{in,i}$ are the system's embeddings for trying to win various games, and the $z_{out,i}$ are the corresponding embeddings for trying to lose those games, then $z_{out}$ could be interpreted as the system's guess at a losing strategy for the game embedded as $z_{in}$, and then could be used to play that game. We could then evaluate performance by how well the system actually performs at losing with the $z_{out}$ strategy. \par

@@

\textbf{Basic meta learning:} First, we show that the system is able to achieve the basic goal of learning a held-out task from a few examples in Fig. \ref{poly_basic_results}. \par

>>>>

\begin{abstract}
How can deep learning systems flexibly reuse their knowledge? Toward this goal, we propose a new class of challenges, and a class of architectures that can solve them. The challenges are in the form of meta-mappings, which involve systematically transforming task behaviors to adapt to new tasks. We suggest that the key to achieving these challenges is representing the task being performed along with the computations used to perform it. We therefore draw inspiration from meta-learning and functional programming to propose a class of Embedded Meta-Learning (EML) architectures that represent both data and tasks in a shared latent space. EML architectures are applicable to any type of machine learning task, including supervised learning and reinforcement learning. We demonstrate the flexibility of these architectures by showing that they can perform meta-mappings, i.e. that they can exhibit zero-shot remapping of behavior on new tasks. 
\end{abstract}

@@

\section{Embedded Meta-Learning (EML) architecture}
To achieve this, we propose a new class of Embedded Meta-Learning (EML) architectures which consist of two parts: 
\vspace{-0.75em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
The utility of having a completely shared space $Z$ in which data, language, and tasks are represented is that it allows for arbitrary mappings between these distinct types of data. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explaining behavior), or tasks to tasks (changing behavior based on examples). This is a step closer to the flexible reasoning humans display. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same data/task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[c.f.][]{Kumaran2016}. \par 
\begin{figure}
\centering
\begin{tikzpicture}[auto, scale=0.75]
\node [dashblock] at (0, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: \textbf{Shared} \\ \textbf{representation}\end{tabular}};

@@

\node [conc] at (-7, -4) (perc) {\textbf{Input}};
%%\node [conc] at (-7.25, -2.5) (lang) {\textbf{Language}};
\node [conc] at (-7, 4) (act) {\textbf{Output}};
\node [block, text width=2cm] at (-4.5, -2) (IE) {$\bm{\mathcal{I}}$: \textbf{Input encoder}};
%%\node [block, text width=2cm] at (-5, -0.5) (LE) {$\bm{\mathcal{L}}$: \textbf{Lang. Encoder}};
\node [block, text width=2.2cm] at (-4, 2.5) (OD) {$\bm{\mathcal{O}}$: \textbf{Output decoder}};
\node [block, text width=2cm] at (-5.5, 1) (TE) {$\bm{\mathcal{T}}$: \textbf{Target encoder}};
\path [line] (perc.north) to (IE.south);
%%\path [line] (lang.north) to ([xshift=0.05cm, yshift=0.05cm]LE.south west);
\path [line] ([xshift=-0.05cm, yshift=-0.05cm]IE.north east) to (rep.south west);
%%\path [line] (LE.east) to (rep.west);
\path [line] (rep.north west) to ([xshift=-0.05cm, yshift=0.05cm]OD.south east);
\path [line] (OD.north) to (act.south);
\path [line] (act.south) to (TE.north);
\path [line] (TE.east) to (rep.north west);

@@

Specifically, we embed inputs, targets, and mappings into a shared representational space $Z$. Inputs are embedded by a deep network $\mathcal{I}: \text{input} \rightarrow Z$. Model outputs are decoded from the representational space by a deep network $\mathcal{O}: Z \rightarrow \text{output}$. Target outputs are encoded by a deep network $\mathcal{T}: \text{targets} \rightarrow Z$. \par
Given this, the task of mapping inputs to outputs can be framed as trying to find a transformation of the representational space that takes the (embedded) inputs from the training set to embeddings that will decode to the target outputs. These transformations are performed by a system with the following components (see Fig. \ref{architecture_fig} for a schematic):  

@@

To write this explicitly, suppose we have some dataset of input, target pairs ($D_1 = \{(x_0, y_0), ...\}$), and some input $x$ from some other dataset $D_2$ for which we wish to generate a predicted output $\hat{y}$. This output would be generated as follows: 

@@

This system can be trained end-to-end if labels are provided for the second set of inputs $D_2$. In particular, suppose we have some loss function $\mathcal{L}(y, \hat{y})$ defined on a single target output $y$ and actual model output $\hat{y}$, for some input $x$. We define our total loss computed on some dataset $D_2$ as:

@@

However, the system is not limited to simply mapping inputs to outputs. The key advantage of our architecture is that it allows transforming anything that is embedded in $Z$ using the same system. For example, suppose we have some task embedding $z_{old} \in Z$, and a meta-mapping embedding $z_{\text{meta}} \in Z$. Then we can generate a new task embedding $\hat{z}_{new} \in Z$ as:  
\[\hat{z}_{new} = f_{z_{\text{function}}}(z_{old}) \qquad \text{where } f_{z_{\text{function}}} \text{ is parameterized by } \mathcal{H}\left(z_{\text{meta}}\right)\]
For example, if the $z_{old}$ is the system's embedding for trying to win at go, and $z_{meta}$ is the meta-task of switching to trying to lose a game, then $z_{new}$ could be interpreted as the system's guess at a losing strategy for go. We could then evaluate performance by how well the system actually loses at go with the $z_{new}$ strategy. \par

@@

\textbf{Basic meta learning:} First, we show that the system is able to achieve the basic goal of learning a held-out polynomial from a few data points in Fig. \ref{poly_basic_results}. \par
